---
title: "Algorithm Development Final Project"
author: "NDR"
date: "January 10, 2016"
output: html_document
---

Dictionary Generation 

To Generate a dictionary for this process, the 1-grams were taken and all the words that appeared more than once were incorporated into our dictionary

```{r}
dict <- read.csv("single_word.csv")
dict <- dict[, 2:3]
dict <- dict[dict$Freq != 1, ]
dict <- dict[, 1]
dict <- as.character(dict)
dict_len <- length(dict)
```

This produces a dictionary with `r dict_len` words in it. The dictionary was used to remove less common words and develop N-grams that take into consideration words that were not found in our training dataset. To do this, the 2-, 3- , and 4-grams data tables were scanned for words not in our dictionary, which were then turned into "<UKN>".  

```{r}
tri <- read.csv("trigram.csv", stringsAsFactors=FALSE)
tri <- tri[,2:5]
for (i in 1:length(tri[,1])) {
        if(!any(tri[i,1] == dict)) {
                tri[i, 1] <- "<UKN>"
        }
        if(!any(tri[i,2] == dict)) {
                tri[i, 2] <- "<UKN>"
        }
        if(!any(tri[i,3] == dict)) {
                tri[i, 3] <- "<UKN>"
        }
}
write.csv(tri, "trigram_Unknown.csv", row.names = F )

duo <- read.csv("duogram.csv", stringsAsFactors=FALSE)
duo <- duo[,2:4]
for (i in 1:length(duo[,1])) {
        if(!any(duo[i,1] == dict)) {
                duo[i, 1] <- "<UKN>"
        }
        if(!any(duo[i,2] == dict)) {
                duo[i, 2] <- "<UKN>"
        }
}
write.csv(duo, "duogram_Unknown.csv", row.names = F )

uni <- read.csv("unigram.csv", stringsAsFactors=FALSE)
uni <- uni[,2:3]
for (i in 1:length(uni[,1])) {
        if(!any(uni[i,1] == dict)) {
                uni[i, 1] <- "<UKN>"
        }
        if(!any(uni[i,2] == dict)) {
                uni[i, 2] <- "<UKN>"
        }
}
write.csv(uni, "unigram_Unknown.csv", row.names = F )
```

Once the n-gram tables were changed to incorporate unknown words, the tables were combined to generate the specified n-grams. For example the first three words (first 3 columns) in the 3-gram table were collapsed into a single character separated by a single space and added as a column to the data tables.   

```{r}
tri$trigram <- paste(tri[,1], tri[,2], tri[,3], sep = " ")
duo$duogram <- paste(duo[,1], duo[,2],  sep = " ")

write.csv(tri, "trigram_Unknown.csv", row.names = F )
write.csv(duo, "duogram_Unknown.csv", row.names = F )
```

```{r}
Tri_results <- data.frame(NULL)
trigrams <- unique(tri$trigram)
for (i in 1:length(trigrams)){
        sub <- tri[tri$trigram == trigrams[i],]
        tab <- as.data.frame(table(sub))
        if (nrow(tab) != 1) {tab <- tab[order(tab$Freq),]}
        Tri_results[i, 1] <- trigrams[i]
        Tri_results[i, 2] <- as.character(tab$outcome[1])
}
colnames(Tri_results) <- c("trigram", "outcome")
write.csv(Tri_results, "trigram_final.csv", row.names = F )
```

```{r}
duo_results <- data.frame(NULL)
duograms <- unique(duo$duogram)
for (i in 1:length(duograms)){
        sub <- duo[duo$duogram == duograms[i],]
        tab <- as.data.frame(table(sub))
        if (nrow(tab) != 1) {tab <- tab[order(tab$Freq),]}
        duo_results[i, 1] <- duograms[i]
        duo_results[i, 2] <- as.character(tab$outcome[1])
}
colnames(duo_results) <- c("duogram", "outcome")
write.csv(duo_results, "duogram_final.csv", row.names = F )
```

```{r}
uni_results <- data.frame(NULL)
unigrams <- unique(uni$var_one)
for (i in 1:length(unigrams)){
        sub <- uni[uni$var_one == unigrams[i],]
        tab <- as.data.frame(table(sub))
        if (nrow(tab) != 1) {tab <- tab[order(tab$Freq),]}
        uni_results[i, 1] <- unigrams[i]
        uni_results[i, 2] <- as.character(tab$outcome[1])
}
colnames(uni_results) <- c("unigram", "outcome")
write.csv(uni_results, "unigram_final.csv", row.names = F )
```